---
title: "CollegeMajorAnalysis"
author: "Shubhendu Dash"
date: "17/07/2020"
output:
  pdf_document: default
  html_document: default
---
# Regression Models:

The goal of the course is use regression analysis as tool to understand your data.

To achieve this objective, the course gives the tools to understand regressions, from the simple and practical linear model, to the glm, focussing on binary outcomes (logit) and count data (poisson).

## Optional Data Analysis Project:

This analysis is part of the optional quiz, to understand how students understand and work with data. 

## A codebook for the dataset is given below:

        rank: Rank by median earnings
        major_code: Major code
        major: Major description
        major_category: Category of major
        total: Total number of people with major
        sample_size: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
        p25th: 25th percentile of earnings
        median: Median earnings of full-time, year-round workers
        p75th: 75th percentile of earnings
        perc_men: % men with major (out of total)
        perc_women: % women with major (out of total)
        perc_employed: % employed (out of total)
        perc_employed_fulltime: % employed 35 hours or more (out of employed)
        perc_employed_parttime: % employed less than 35 hours (out of employed)
        perc_employed_fulltime_yearround: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
        perc_unemployed: % unemployed (out of employed)
        perc_college_jobs: % with job requiring a college degree (out of employed)
        perc_non_college_jobs: % with job not requiring a college degree (out of employed)
        perc_low_wage_jobs: % in low-wage service jobs (out of total)
        
## Data Analysis Process Starts:

1. Installing required packages.

        1.1 To get started, start a new R/RStudio session with a clean workspace. To do this in R, you can use the q() function to quit, then reopen R. The easiest way to do this in RStudio is to quit RStudio entirely and reopen it. After you have started a new session, run the following commands. This will load a data.frame called college for you to work with.

```{r echo=TRUE, cache=TRUE}
#This will load a data.frame called college for you to work with.

install.packages("devtools")
devtools::install_github("jhudsl/collegeIncome", force = TRUE)
library(collegeIncome)
data(college)

#Next download and install the matahari R package with the following commands.

devtools::install_github("jhudsl/matahari", force = TRUE)
library(matahari)
```

        1.2 This package allows a record of your analysis (your R command history) to be documented. You will be uploading a file containing this record to GitHub and submitting the link as part of this quiz.

        1.3 Before you start the analysis for this assignment, enter the following command to begin the documentation of your analysis:

```{r echo=TRUE}
dance_start(value = FALSE, contents = FALSE)
```

2. Calling required packages from the library.

```{r echo=TRUE}
library(ggplot2)
library(dplyr)
library(GGally)
```

3. Getting into the Analysis.

```{r echo=TRUE}
#Initial phase of data.frame analysis.

dim(college)
str(college)
```

        3.1 How's the distribution of fulltime employment? Is there a gender difference?

```{r echo=TRUE}
g <- ggplot(aes(perc_employed_fulltime_yearround, major_category, colour=perc_women), data=college)
g <- g + geom_point(size=2) + labs(x="Percentage Employed Fulltime Yearround", y="Major Category")
g <- g + scale_colour_gradient(name="Percentage of Women")
```

        3.2 Is there a difference in the median of earnings?
        
```{r echo=TRUE}
h <- ggplot(aes(median, major_category, colour=perc_women), data=college)
h <- h + geom_point(size=2) + labs(x="Median earnings of full-time, year-round workers", y="Major Category") + scale_colour_gradient(name="Percentage of Women")
ggmatrix(list(g, h), ncol=2, nrow=1, legend=grab_legend(g), xAxisLabels = c("Percentage Employed Fulltime Yearround","Median earnings of full-time, year-round workers" ))
```

        3.3 Looking median, 25th and 75th quintile of earnings.
        
```{r echo=TRUE}
p25th <- data.frame(Earnings = college$p25th, Major_category = college$major_category, Major = college$major, Value="25th Quantile")
median <- data.frame(Earnings = college$median, Major_category = college$major_category, Major = college$major, Value="Median")
p75th <- data.frame(Earnings = college$p75th, Major_category = college$major_category, Major = college$major, Value="75th Quantile")
earnings <- rbind(p25th, median, p75th)
m <- ggplot(aes(Earnings, Major_category), data=earnings)
m <- m + geom_point(aes(colour=Value)) + labs(x="Earnings of full-time, year-round workers", y="Major Category")
m #The higher earnings are Engineering, Education, Business and Biology & Life Science
```

        3.4 Investigating these majors further.

```{r echo=TRUE}
big4 <- filter(earnings, Major_category=="Engineering"| Major_category=="Education"| Major_category=="Business" | Major_category=="Biology & Life Science")
l <- ggplot(aes(Earnings, Major), data = big4) + facet_grid(Major_category ~ .)
l <- l + geom_point(aes(colour = Value)) + labs(x="Earnings of full-time, year-round workers", y=NULL)
l
```

        3.5 Looking at all the correlations.

```{r echo=TRUE}
all <- ggpairs(data = select(college, c(major_category, total, perc_women, median, perc_employed_fulltime:perc_low_wage_jobs)), cardinality_threshold = 17)
all
```

        3.6 Regressions.

```{r echo=TRUE}
fit <- lm(median ~ factor(major_category), data=college)
summary(fit)

#The major categories weren't significant

fit2 <- lm(median ~ factor(major_category) + perc_women, data = college)
anova(fit, fit2)

#Still not able to reject the null hypothesis, thus the regressors are not significant

fit3 <- lm(median ~ perc_women, data = college)
anova(fit3, fit2)
fit4 <- lm(median ~ factor(major_category)*perc_women, data=college)
summary(fit4)

#Now there are some very significant coef
anova(fit, fit4)

#The inclusion of the interation between the percentage of woman and the major category

anova(fit3, fit4) 

#was able to reject the null hypothesis, thus indicating the significance
#Still, the R squared is low, even though the large number of variables, as shown by the adjusted R-squared

fit5 <- lm(p25th ~ factor(major_category)*perc_women, data=college)
summary(fit5)

#The results for the model using the 25th quintile don't hold as well as the model for the median earnings
#The R^2 is lower, and it's unable to reject the null hypothesis for the F-statistic 

fit6 <- lm(p75th ~ factor(major_category)*perc_women, data=college)
summary(fit6)

#The model for the 75th quintile has a smaller adjusted p-value
```

        3.7 Conclusions.
        
- An interesting conclusion is that for 3 out of the 4 major categories investigated further, 
- Education, Engineering, Business and Biology & Life Science, because of the high earnings,their interations with the percentage of women in these areas were significant and their coefficients were negative. Looking at the graphic "h" it's possible to visualize this effect.
- Even though this major categories appear to receive higher earnings, the greater the female participation in the area,the lower their gains.

        3.8 Looking at the residuals.

```{r echo=TRUE}
par(mfrow=c(1,2))
plot(resid(fit), pch=16, col="deepskyblue4")
title("Residuals from median ~ major category")
abline(h=0, col="brown3", lwd=2)
plot(resid(fit4), pch=16, col="deepskyblue4")
abline(h=0, col="brown3", lwd=2)
title("Residuals from median ~ major category * percetage of women")

#The residuals for the model with the interation are closer to zero, and appear random enough.

fit7 <- lm(median ~ major_category*perc_women + perc_college_jobs + perc_employed + perc_low_wage_jobs, data = college)
summary(fit7)
```

```{r echo=TRUE}
#When you have finished your analysis, use the following command to save the record of your analysis on your desktop:

dance_save("college_major_analysis.rds")
```

## Conclusion:

- Including other variables doesn't bring much to the model, their coefficients are not significant and the R^2 (and adjusted R^2) and the F-statistic are slightly worse.
- The best model was the forth, with the interaction between the percentage of women and the major category.
- Some interesting coefficients are significant and the model and its F-statistic is able to reject the null hypothesis that all the coefficients are zero. But still, the effect is minimal, and the model doesn't explain much of the outcome.
- The first model, with only the major categories as regressors indicate that it does not affect the earnings.
